### 2024-03-06 README/getting_started_guide.tex

% !TeX root = README.tex

\section{Getting Started Guide}

Our artifact, dubbed Hybrid Resource-Aware ML (RaML), is a program analysis tool
that takes in an OCaml program and infers its worst-case polynomial cost bound
using the technique Hybrid AARA described in the paper.

The artifact is wrapped inside the provided Docker image.
%
To run it, first install Docker according to the instructions at
\url{https://docs.docker.com/engine/install/}.
%
To see if Docker has been installed properly, run
\begin{verbatim}
$ docker --version
Docker version 25.0.3, build 4debf41
\end{verbatim}

Load and run the Docker image as follows:
\begin{verbatim}
$ docker load --input hybrid_aara.tar.gz
$ docker run -it --rm hybrid_aara
root@fca41f6f696d:/home/hybrid_aara/benchmark_suite/playground#
\end{verbatim}

In the initial working directory, we have a file \texttt{append.ml} that stores
the \texttt{append} function written in OCaml.
%
Given two inputs lists, the \texttt{append} function appends the first input to
the second one.
%
For each element in the first input that the function processes, it incurs a
cost ranging from 0.5 to 1.0.
%
Therefore, the worst-case cost bound of the function is $1.0 \cdot M$, where $M$
is the length of the first input list.

Hybrid RaML's executable is the file \texttt{/home/hybrid\_aara/raml/main}, and
its location has already been added to the \texttt{PATH} variable so that we can
call Hybrid RaML by running \texttt{main} anywhere in the filesystem.
%
Hybrid AARA (and its artifact Hybrid RaML) provides three types of hybrid
resource analysis: Hybrid \Opt{}, Hybrid \BayesWC{}, and Hybrid \BayesPC{}.

\subsection{Hybrid \Opt{}}

To perform Hybrid \Opt{} on the \texttt{append} function, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_opt.json
\end{verbatim}
%
In the command, \texttt{stat\_analyze} means we want to perform Hybrid AARA
(instead of the conventional AARA).
%
The input \texttt{ticks} means the resource metric of our interest is the tick
metric (instead of heap memory, for example).
%
The number 1 means we want to infer degree-1 polynomial cost bounds.
%
The input \texttt{append.ml} specifies the input file name, and the input
\texttt{append} means we want to know a cost bound of the function
\texttt{append} (instead of other functions in the same file).
%
Finally, the input \texttt{config\_opt.json} specifies the configuration file
for Hybrid \Opt{}.

In the output printed on the stdout, you should find
\begin{verbatim}
== append :

[int list; int list] -> int list

Non-zero annotations of the argument:
    0.85  <--  ([::(*)], [])

Non-zero annotations of result:

Simplified bound:
    0.85*M
  where
    M is the number of ::-nodes of the 1st component of the argument
\end{verbatim}
%
It means that the cost bound inferred by Hybrid \Opt{} is $0.85 \cdot M$, where
$M$ is the first input's length.

\subsection{Hybrid \BayesWC{}}

Next, to perform Hybrid \BayesWC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayeswc.json
\end{verbatim}
where the configuration file is now \texttt{config\_bayeswc.json}, which
contains hyperparameters for Bayesian inference.
%
The analysis should finish in between 40 and 50 seconds.
%
Its output on the stdout should contain
\begin{verbatim}
== append :

  [int list; int list] -> int list

  Non-zero annotations of the argument:
(avg, std) = (   2.18047,   0.633256)  <--
                                         ([::(*)], [])
(avg, std) = ( 0.0101388,  0.0158341)  <--  ([], [::(*)])
(avg, std) = (         0,          0)  <--  ([], [])

  Non-zero annotations of result:
(avg, std) = (         0,          0)  <--
                                   [::(*)]
(avg, std) = (         0,          0)  <--  []

  Simplified bound:
     0.0101388*L + 2.18047*M
   where
     L is the number of ::-nodes of the 2nd component of the argument
     M is the number of ::-nodes of the 1st component of the argument
\end{verbatim}
%
Unlike Hybrid Opt{} that infers a single cost bound, Hybrid \BayesWC{} returns a
collection of cost bounds, which approximate their posterior distribution.
%
In the stdout, 0.0101388 is the average degree-one coefficient for the first
input in the posterior distribution, and likewise, 2.18047 is the average
coefficient for the second input.

\subsection{Hybrid \BayesPC{}}

To perform Hybrid \BayesPC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayespc.json
\end{verbatim}
using the configuration file \texttt{config\_bayespc.json}.
%
The analysis should terminate within 90 seconds.
%
The output printed on the stdout should contain
\begin{verbatim}
== append :

  [int list; int list] -> int list

  Non-zero annotations of the argument:
(avg, std) = (   2.03295,   0.228391)  <--  
                                         ([::(*)], [])
(avg, std) = (  0.281939,   0.146543)  <--  ([], [::(*)])
(avg, std) = (  0.164309,  0.0824471)  <--  ([], [])

  Non-zero annotations of result:
(avg, std) = (         0,          0)  <--  
                                   [::(*)]
(avg, std) = (         0,          0)  <--  []

  Simplified bound:
     0.164309 + 0.281939*L + 2.03295*M
   where
     L is the number of ::-nodes of the 2nd component of the argument
     M is the number of ::-nodes of the 1st component of the argument
\end{verbatim}
where \texttt{0.164309 + 0.281939*L + 2.03295*M} contains the average
coefficients.