### 2024-03-06 README/step_by_step_instructions.tex

% !TeX root = README.tex

\section{Step-by-Step Instructions}