### 2024-03-07 README/step_by_step_instructions.tex
Fleshed out the explanation of how to produce plots and tables

% !TeX root = README.tex

\section{Step-by-Step Instructions}

\subsection{Benchmark Suite}

The artifact contains 10 benchmark programs for Hybrid AARA.
%
As listed in Section 7 of the paper, these benchmarks are: \texttt{MapAppend},
\texttt{Concat}, \texttt{InsertionSort2}, \texttt{QuickSort},
\texttt{QuickSelect}, \texttt{MedianOfMedians}, \texttt{ZAlgorithm},
\texttt{BubbleSort}, \texttt{Round}, and \texttt{EvenOddTail}.
%
Inside the artifact, the benchmark \texttt{EvenOddTail} goes by its full name,
\texttt{EvenSplitOddTail}, instead of its shorthand \texttt{EvenOddTail} used
throughout the paper.
%
We have tested both data-driven resource analysis and hybrid resource analysis
on the first seven benchmarks, and only data-driven resource analysis on the
remaining three benchmarks.

Inside the benchmark directory \texttt{/home/hybrid\_aara/benchmark\_suite},
each benchmark has its own directory.
%
Furthermore, inside the benchmarks' directories, we have the following
directories:
\begin{enumerate}
  \item \texttt{utility}: Python source files that contain (i)
        benchmark-specific configurations for hybrid resource analysis and (ii)
        functions for analyzing the inference result.
  \item \texttt{bin}: input and output files for Hybrid RaML.
  \item \texttt{images}: plots of inferred cost bounds.
\end{enumerate}

\subsection{Display Experiment Results}

In this section, we demonstrate how to create tables and plots in the paper from
the experiment results that are stored in each benchmark's \texttt{bin}
directories.
%
In the Docker container, the directory \texttt{bin} of each benchmark already
contains the experiment results that we reported in the paper.
%
We will later demonstrate how to generate the experiment results from scratch in
\cref{sec:generate experiment results}.

\paragraph{Table~1}

In Table~1 in the paper, the 4\textsuperscript{th} and 5\textsuperscript{th}
columns display the percentages of sound cost bounds inferred by data-driven and
hybrid resource analyses, respectively.
%
To display these two columns, run
\begin{verbatim}
# cd /home/hybrid_aara/benchmark_suite/toolbox
# python3 run_analysis.py soundness
\end{verbatim}

The 6\textsuperscript{th} and 7\textsuperscript{th} columns of Table~1 display
the analysis time of data-driven hybrid resource analyses, respectively.
To display these two columns, run
\begin{verbatim}
# cd /home/hybrid_aara/benchmark_suite/toolbox
# python3 run_analysis.py time
\end{verbatim}

\paragraph{Figure~5}

Figure~5 in the paper displays relative errors of 5 selected benchmarks.
%
To create this plot, run
\begin{verbatim}
# cd /home/hybrid_aara/benchmark_suite/toolbox
# python3 run_analysis.py plot_relative_errors
\end{verbatim}
%
This command creates a plot and stores it as a file
\texttt{relative\_errors.pdf} inside the directory
\texttt{/home/hybrid\_aara/benchmark\_suite/images}.
%
To view the plot, we first transfer it from the Docker container to your local
filesystem.
%
Run the following command in the terminal of your local machine:
\begin{verbatim}
$ docker cp hybrid_aara:/home/hybrid_aara/benchmark_suite\
/images/relative_errors.pdf \
<path-to-local-directory>
\end{verbatim}
where the second argument \texttt{<path-to-local-directory>} is a desired
destination directory on your local filesystem.
%
You can then view the plot on your local machine.

\paragraph{Figures~6--24}

For Figures~6--24 in the paper, to create them, run
\begin{verbatim}
# cd /home/hybrid_aara/benchmark_suite/toolbox
# python3 run_analysis.py plot
\end{verbatim}
%
This command creates all plots of inference results and stores them in the
directory \texttt{benchmark\_suite/<benchmark-name>/images}.
%
For example, for hybrid BayesWC performed on \texttt{QuickSort}, we can
find its plot in the file \texttt{inferred\_cost\_bound.pdf} in the directory
\texttt{benchmark\_suite/QuickSort/images/hybrid/bayeswc}.

To view the plot, again, we transfer it from the Docker container to your local
filesystem.
%
Run this command on the terminal of the local machine:
\begin{verbatim}
$ docker cp hybrid_aara:/home/hybrid_aara/benchmark_suite\
/QuickSort/images/hybrid/bayeswc/inferred_cost_bound.pdf \
<path-to-local-directory>
\end{verbatim}
%
You can then view the plot on your local machine.

\paragraph{Tables~2--11}

For Tables~2--11 in the paper, to create and display them on the stdout, run
\begin{verbatim}
# cd /home/hybrid_aara/benchmark_suite/toolbox
# python3 run_analysis.py relative_errors
\end{verbatim}

\subsection{Generate Experiment Results}
\label{sec:generate experiment results}

\subsection{Source Code}