### 2024-03-08 README/what_is_included.tex
Added a list of items necessary for artifact evaluation reviewers

% !TeX root = README.tex

\section{What is Included}

From Zenodo, you should obtain the following items:
\begin{itemize}
  \item This \texttt{README.pdf} document
  \item Archived and compressed Docker image \texttt{hybrid\_aara.tar.gz}
  \item Update version of the paper \texttt{paper.pdf}.
\end{itemize}