### 2024-03-09 README/what_is_included.tex

% !TeX root = README.tex

\section{What is Included}

From Zenodo, you should obtain the following items:
\begin{itemize}
  \item This \texttt{README.pdf} document
  \item Archived and compressed Docker image \texttt{hybrid\_aara.tar.gz}
  \item Updated version of the paper \texttt{paper.pdf}.
\end{itemize}

\emph{\textbf{Note:} The tables and plots produced by the artifact are
  slightly different from the ones in the submitted version of the
  paper.
%
  This is because, since the submission, we have adjusted the
  hyperparameters for \BayesWC{} and \BayesPC{} using an automatic
  procedure for choosing hyperparameters.
%
  The updated experiments results still preserve all key findings in
  Section~7 of the paper.
%
  The updated version of the paper accompanies the artifact.}