### 2024-03-11 README/what_is_included.tex
Minor revision

% !TeX root = README.tex

\section{What is Included}

From Zenodo, you should obtain the following items:
\begin{itemize}
  \item This \texttt{README.pdf} document
  \item Archived and compressed Docker image \texttt{hybrid\_aara.tar.gz}
  \item Updated version of the paper \texttt{paper.pdf}.
\end{itemize}

\emph{\textbf{Note:} The tables and plots produced by the artifact are slightly
  different from the ones in the submitted version of the paper.
  %
  This is because, since the submission, we have adjusted the hyperparameters
  for \BayesWC{} and \BayesPC{} using an automatic procedure for choosing
  hyperparameters (Appendix~C of the paper), as requested by the paper
  reviewers.
  %
  The updated experiments results still preserve all key findings in Section~7
  of the paper.
  %
  The updated version of the paper accompanies the artifact.}