### 2024-03-29 reviews/response_reound1.md
Added a review and response

Thank you all for the feedback.

# Review A

Thank you for taking the time to run all Python scripts. We are happy that you
did not encounter any technical issues with the artifact.

# Review B

Thank you for going through the getting-started guide. We are happy that you did
not encounter any technical issues with the artifact.

# Review C

> The binary included in the docker image crashes immediately (Illegal
> instruction (core dumped)) on my architecture (described below to ease
> debugging).

Thank you for reporting this issue. We suspect that the Docker container
requires an instruction that your computer does not support. We built the docker
image on the OS Ubuntu 22.04 LTS and the CPU Intel(R) Core(TM) i7-7600U CPU @
2.80GHz.

We could not identify the root cause of the issue. So could you please run the
artifact by cloning the GitHub repository and building a Docker image locally on
your machine? We are happy that at least you successfully built the image on
your machine.

> It would be good to add instructions in the README to recompile the code
> directly inside the docker, and avoid those crashes. This would additionally
> allow user to tweak the code directly inside the docker, and observe changes
> there.

Thank you for the suggestion. Technically, from the Dockerfile in the GitHub
repository, one can learn how to recompile the code inside the Docker container.

In the revision of `README.pdf`, we will provide details on how to recompile (i)
the C++ code of the volesti-RaML interface and (i) the OCaml code of Hybrid
RaML. The remaining code, such as the Python scripts, does not need
recompilation after modification.

> Being able to test to analyser on new benchmarks is required to grant the
> "Reusable" badge. It would be helpful to add a small paragraph in the README
> instructing how to do so, and include tools in the docker image to make this
> process easier.

If the user wishes to analyze their custom OCaml program using Hybrid RaML, the
user should first prepare (i) the source file of their OCaml program, augmented
with a list of inputs for runtime cost data generation and (ii) a configuration
file specifying hyperparameters of Hybrid AARA. To learn the formats of these
files, the user can look at the examples inside the artifact. Finally, The user
runs Hybrid RaML as demonstrated in Section 1 of `README.pdf`.

We will explain this point in the revision of `README.pdf`.

> It would improve the artifact to include an editor.

We will install the text editor Vim in the updated Docker image. Alternatively,
if the user wishes to install a different text editor, while the Docker
container is running, the user can run `apt update && apt install <editor-name>
-y` inside the container.

> Should input_dataset be entered manually by the users on each new benchmark?
> Are scripts provided to generate them automatically?

The artifact already comes with a Python script (`input_data_generation.py`
inside the directory `home/hybrid_aara/benchmark_suite/toolbox/`) for generating
lists, pairs of lists (for the benchmark MapAppend), and nested lists (for the
benchmark Concat). For example, the function `create_exponential_lists` inside
this script generates lists whose sizes grow exponentially (e.g., 1, 2, 4, 8,
16, etc.). Here, the initial size (i.e., 1) and exponent (i.e., 2) are specified
by the user. The content of each list is determined randomly. The function
`convert_lists_python_to_ocaml` then converts a list from the Python format to
the OCaml format.

We will mention this Python script in the revised `README.pdf`. Also, we will
revise the script such that, when the user runs `python3
input_data_generation.py`, the script prints out an example collection of input
lists used for runtime cost data generation.